Tsbrowse: a genome browser for Ancestral Recombination Graphs

### 2025-03-27 manuscript/paper.tex
Good opening para for Applications

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

@@

\subsection{Applications}

 Tsbrowse can be applied to versatile scenarios. Figure \ref{fig:Figure_2}
highlights a use-case in visually characterising ARGs. The input ARG is a
selective sweep dataset simulated using \texttt{msprime}
\citep{baumdicker2022efficient}. The distinct pattern of edges observed in the middle of
the sequence between \textit{y=0} to \textit{y=2.5} is attributable to a burst
in recent coalescent events, showing that tsbrowse captures this characteristic
of a selective sweep. Supplementary Figure \ref{fig:Supplementary_Figure_1} is
another example of characterising ARGs with tsbrowse. The view of edges from
four different inferences of the same dataset (inferred with \texttt{tsinfer} +
\texttt{tsdate} (Panel A), \texttt{Relate} (Panel B), \texttt{ARG-Needle}
(Panel C) and \texttt{SINGER} (Panel D)), reflects qualitative differences
between inference methods.

>>>>

%Fig~\ref{fig:Figure_2} %
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

@@

\subsection{Applications}
The purpose of \texttt{tsbrowse} is to provide an interactive view of the 
\texttt{tskit} ARG data model to guide intuition, improve inference quality control
and facilitate debugging. An example of how we can deepen our understanding
using the genome-browser-like perspective provided by \texttt{tsbrowse}
is provided the simulated ARG of Fig~\ref{fig:Figure_2}, where 
the gap in the edges view corresponds to the characteristic dip in diversity
of a selective sweep. Comparing this ground truth to the ARGs inferred
by four different inference methods in Fig~\ref{fig:Supplementary_Figure_1},
we can see that there are substantial qualitative differences between the
results. These differences in ancestral haplotypes illustrated by 
the edges view are unlikely to be captured by the tree-by-tree distance
metrics usually used to evaluate
inference methods~\citep[e.g.][]{kelleher2019inferring,zhang2023biobank},
providing further motivation for new and improved ways to compare
simulated and inferred ARGs~\citep{fritze2024forest}.
